Clonal interference between aneuploidy and mutation

### 2021-06-07 ms/ms.tex
likelihood functions updated

@@

\paragraph{Data.} We inference parameters of the models based on the data of the experiments of Yona at al ~\cite{Yona2012}. In \emph{heat-stress} experiment four populations of \emph{S. cerevisiae} were evolved under $39^{\circ}$ heat-stress and developed aneuploidy till 450 generations (here and further by development/loss of genotype trait we mean that 95\% of the population are the carriers of the genotype). Then, two populations were checked at generations 1700 and 2350, and for one population the aneuploidy were eliminated and has not observed allready at generation 1700 and for the second at generation 2350. \\
Similarly, in \emph{ph-stress} experiment four populations were evolved unde high-ph stress conditions and developed aneuploidy during 150 generations. At generation 750 no aneuploidy was observed for all four populations.

\paragraph{Likelihood function.}
 Denote by $A_{450}$ the event of aneuploidy observation at generation 450; by $L_{1700}$ the event of aneuploidy loss observation at generation 1700; and by $L_{2350^*}$ the event of aneuploidy loss observation at generation 2350 while there's no loss in generation 1700. The likelihood function of the model with parameter set $\phi$ for \emph{heat-stress} experiment is:
 \begin{multline} \label{eq:heatstress-likelihood}
L(\phi | A_{450}, L_{1700}, L_{2350^*} ) = \\
P(L_{450})^4 \cdot (1 - P(\lnot{L_{1700}} | A_{450} )^4 - P(\lnot{L_{2350^*}}| A_{450})^4 + P(\lnot{L_{1700}} \land \lnot{L_{2350^*}}| A_{450})^4 )		
 \end{multline}

>>>>

% TODO cite https://www.biorxiv.org/content/10.1101/2020.10.07.327759v1

@@

\paragraph{Data.} We inference parameters of the models based on the data of the experiments of Yona at al ~\cite{Yona2012}. In \emph{heat-stress} experiment four populations of \emph{S. cerevisiae} were evolved under $39^{\circ}$ heat-stress and developed aneuploidy till 450 generations (here and further by development/loss of genotype trait we mean that 95\% of the population are the carriers of the genotype). Then, two populations were checked at generations 1700 and 2350, and for one population the aneuploidy were eliminated and has not observed allready at generation 1700 and for the second at generation 2350. \\
Similarly, in \emph{ph-stress} experiment four populations were evolved unde high-ph stress conditions and developed aneuploidy during 150 generations. At generation 750 aneuploidy fully eliminated for two populations, and only partly for another two.

\paragraph{Likelihood function.}
Denote by $A_{t}$ the event of aneuploidy observation at generation $t$; by $L_{t}$ the event of aneuploidy loss observation at generation $t$; and by $L_{2350^*}$ the event of aneuploidy loss observation at generation 2350 while there's no loss in generation 1700. The likelihood function of the model with parameter set $\theta$ for \emph{heat-stress} experiment is:
\begin{multline} \label{eq:heatstress-likelihood}
L(\theta | A_{450}, L_{1700}, L_{2350^*} ) = \\
P^4(A_{450}) \cdot (1 - P^4(\lnot{L_{1700}} | A_{450} ) - P^4(\lnot{L_{2350^*}}| A_{450}) + P^4(\lnot{L_{1700}} \land \lnot{L_{2350^*}}| A_{450}) )		
\end{multline}
The likelihood function of the model with parameter set $\theta$ for \emph{ph-stress} experiment is:
\begin{equation} \label{eq:phstress-likelihood}
 L(\theta | A_{150}, L_{750} ) = \\
 P^4(L_{450}) \cdot 6 \cdot	P^4(L_{750} | A_{150}) \cdot P^4(\lnot{L_{750}} | A_{150})
\end{equation}

\paragraph{Parameter inference.} To estimate the parameters of the models we apply approximate Bayesian computation with a sequential Monte-Carlo scheme (ABC-SMC), employing pyabc Python 
package \citep{Klinger2018}
